---
title: "EDA_ProductAnalysis"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
mytheme <- theme(panel.grid.major = element_blank()
                 , panel.grid.minor = element_blank()
                 ,panel.background = element_rect(fill = '#1b212c')
                 ,plot.background  = element_rect(fill = '#1b212c', colour = 'white')
                 ,axis.text = element_text(family = NULL, face = "bold", colour = "white",size = 8)
                 ,axis.title = element_text(family = NULL, face = "bold", colour = "white",size = 10)
                 ,legend.background = element_rect(fill = '#1b212c', colour = 'white'))
```

```{r cars, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(ggplotify)

df <- read.csv("./../DataExtract/Data/processed.json_partial.csv")
df <- as.data.frame(df)
print(colnames(df))
```

### Data Correction

```{r }
df$totalRatingsNum <- as.numeric(gsub("," , "", as.character(df$totalRatings)))
df$rank <- as.numeric(gsub("," , "", as.character(df$rank)))
```
### Correcting Brand Names


```{r pressure}
table(df$brand)

# correcting l'oreal spellings
df[which(regexpr("L'" , df$brand) >= 0), ]$brand <- "L'Oreal"
```

### Price Analysis

``` {r }

## Estimated bands
df %>% 
  group_by(brand) %>%
  summarise(meanprice = mean(price))

## Price
ggplot(data = df , aes(x = df$price)) + 
  geom_density(color = "white") + 
  geom_vline(xintercept = 8.45 , color="purple" ) + 
  mytheme

df$priceband <- case_when( df$price < 7.5 ~ "Low"
                           , df$price < 12.5 ~ "Medium"
                           ,TRUE ~ "High")
```


### Ratings Analysis

```{r }
## Ratings
ggplot(data = df , aes(x = df$stars)) + 
  geom_density() + 
  scale_x_continuous(breaks = seq(2.5,  5 , 0.1))

df$starsBand <- case_when( df$stars < 3.9 ~ "Poor"
                          ,df$stars < 4.3 ~ "Good"
                          ,TRUE ~ "Excellent")
```


### Price Vs Ratings

``` {r }
table(df$starsBand , df$priceband)

ggplot(data = df, aes(x = df$stars , y = df$price)) + geom_jitter(width = 0.1)
```

```{r }



ggplot(data = df, aes(x = df$price , y = df$totalRatingsNum)) + geom_jitter(width = 0.1)

ggplot(data = df, aes(x = df$totalRatingsNum)) + geom_density()
ggplot(data = df, aes(x = log(df$totalRatingsNum))) + geom_density()

ggplot(data = df, aes(x = df$price , y = log(df$totalRatingsNum))) + geom_jitter(width = 0.1 , color="white") + geom_smooth() + mytheme
```

There is a decreasing trend in the number of ratings as the price increase, so fewer people buy products with higher price.


### Demand Curve

```{r}

baseprice <- 4
demandDf <- df[,c("totalRatingsNum" , "price")] %>% arrange(desc(price))
demandDf$demand <- cumsum(demandDf$totalRatingsNum)
demandDf$profit <- (demandDf$price - baseprice)*demandDf$demand
head(demandDf)

ggplot(data = demandDf , aes(x = price , y = profit)) +
  geom_point(color = "white") +
  geom_smooth() + mytheme
```
### Amazon Rank Vs Price , Rating , Total Ratings

```{r}

ggplot(data = df , aes(x =df$price , y=log(df$rank) )) + geom_jitter(width = 0.1) + geom_smooth()
ggplot(data = df , aes(x =df$stars , y=log(df$rank) )) + geom_jitter(width = 0.1) + geom_smooth()
ggplot(data = df , aes(x =log(df$totalRatingsNum) , y=log(df$rank) )) +
  geom_jitter(width = 0.1 , color="white") + 
  geom_smooth() + mytheme
```

As price increases the products are not as popular. Rank decreases as the number of people rating the product increases.


### Color Analysis

* Color Distribution

``` {r }
colorDf <- read.csv("./../DataExtract/Data/processed.json_color.csv")
colorDf$Color <- as.character(colorDf$Color)
splitColors <- strsplit(colorDf$Color , split = ' ')

combinedColors <- Reduce(f = function(a,b) c(a , b) , x = splitColors , init=list())
freqColors <- as.data.frame(table(unlist(splitColors)))
freqColors$Var1 <- as.character(freqColors$Var1)

sortedFreqColors <- freqColors %>% arrange(desc(Freq))
sortedFreqColors$Var1 <- factor(as.character(sortedFreqColors$Var1), levels=sortedFreqColors$Var1 )

head(sortedFreqColors)

ggplot(data= sortedFreqColors[1:20,] , aes(x = Var1 , y=Freq)) +
  theme(axis.text=element_text(size=6)) +
  geom_bar(stat='identity') + 
  coord_flip() + mytheme

```

* Summarising Various Colors

```{r}
#combined = data.frame(price = rep(colorDf[,c("price", "totalRatings")], sapply(splitColors, length)), color = unlist(splitColors))

repd <- colorDf[rep(row.names(colorDf) ,sapply(splitColors, length) ) , c("stars" ,"price" , "totalRatings" , "rank" , "brand")]
repd$color <- unlist(splitColors)

repd$totalRatingsNum <- as.numeric(gsub("," , "", as.character(repd$totalRatings)))
repd$rank <- as.numeric(gsub("," , "", as.character(repd$rank)))
repd[which(regexpr("L'" , repd$brand) >= 0), ]$brand <- "L'Oreal"

head(repd)

meanByColor <- repd %>% 
  group_by(color) %>%
  summarise(meanStar=mean(stars), meanPrice = mean(price) , meanNumRatings = mean(totalRatingsNum) , meanRank = mean(rank) , totalCount = n()) %>%
  arrange(desc(meanNumRatings))

head(meanByColor)

```

```{r}


ggplot(data = meanByColor , aes(x = meanPrice , y= meanRank , label=color , color = color)) +
    geom_jitter(width=1 , show.legend = FALSE) +
    geom_text(alpha = 0.4, show.legend = FALSE)


ggplot(data = meanByColor , aes(x = meanPrice , y= meanRank , label=color , color = color)) +
    geom_text(alpha = 0.6, show.legend = FALSE) + 
  scale_x_continuous(limits = c(5,10)) + 
  scale_y_continuous(limits = c(0,10000)) 


library(treemap)

repd %>% group_by(brand ,color) %>% summarise(mprice = mean(price))


# Garnier is most popular. Popluarity proportional to total Number of ratings
treemap(repd,
        vSize = "totalRatingsNum",
        vColor = "price",
        index = c("brand" , "color")
        ,palette = "Blues"
        , fun.aggregate = "mean"
        , position.legend = "right"
        , type="value")


# Brown and Blonde are Ranked by amazon at the top in most brands
# Got2b is struggling as people don't really like their products
repd$rankInverse = 1/repd$rank
treemap(repd,
        vSize = "rankInverse",
        vColor = "stars",
        index = c("brand" , "color")
        ,palette = "Spectral"
        , fun.aggregate = "mean"
        , position.legend = "right"
        , type="value")


# RAW is being used by fewer people but they seem to like it
# It has very different palette on offer compared to others. Very stark colors.
# Most colors in this chart are ones that are less common.
repd$numRatingInverse = 1/repd$totalRatingsNum
treemap(repd,
        vSize = "numRatingInverse",
        vColor = "stars",
        index = c("brand" , "color")
        ,palette = "Spectral"
        , fun.aggregate = "mean"
        , position.legend = "right"
        , type="value")
```



